---
title: "Linseed Tutorial"
author: "Konstantin Zaitsev"
date: "March 13, 2018"
output: html_document
---
# Linseed tutorial

Linseed (LINear Subspace identification for gene Expresion Deconvolution) is a package that provides tools and interface to explore gene expression datasets in linear space.

## Load the library

```{r warning=FALSE, message=FALSE}
library(linseed)
```

## Getting started with linseed

To start working with gene expression data, we need to create a new LinseedObject, in this tutorial we will use GSE19830 (mixture of Liver, Brain and Lung), we will take only mixed samples (10-42) and will take only 10000 most expressed genes.

```{r warning=FALSE, message=FALSE}
lo <- LinseedObject$new("GSE19830", samples=10:42, topGenes=10000)
```

## Coolinearity networks

To build a coolinearity network we first have to evaluate all pairwise collinearity coefficients and then select genes that have at least one gene that is very collinear ( $R^2 >= 0.99$ ) with it and 10 genes that are less collinear ( $R^2 >= 0.95$ ). We only leave genes that meet both requirements. Be very carefull `calculatePairwiseLinearity` is very memory-demanding method ( $N^2$ memory where $N$ is the number of genes).

```{r networks, }
lo$calculatePairwiseLinearity(negToZero=T)
genes1 <- lo$getGenesByCutOff(0.99, 1)
genes2 <- lo$getGenesByCutOff(0.95, 10)
goodGenes <- intersect(genes1, genes2)
subnetwork <- lo$pairwise[goodGenes, goodGenes]

```

To visualiaze network we can use pheatmap package or use `hclust` itself.

```{r visi}
library(pheatmap)
library(RColorBrewer)

pheatmap(subnetwork, clustering_method = "average",
         show_rownames = F, show_colnames = F, 
         color = colorRampPalette(c("blue", "white", "red"))(100),
         border_color = NA)

```

## Complete deconvolution

First step is identification of the number of the cell types presented in the mixture. 

```{r warning=FALSE, message=FALSE}
lo$svdPlot(dataset="norm")
```

we can suggest from the figure, that since most of the variance is explained by first three singular vectors of SVD, that our dataset is as mixture of 3 cell types.

## Projection

Once we know number of cell types, we can project the data into a plane.

```{r warning=FALSE, message=FALSE}
lo$setCellTypeNumber(3)
lo$project()
lo$projectionPlot()
```


## Corner identification

Once we projected the data into plane, we can identify simplex corner using SISAL algorithm and then select genes closest to identified corners. 

```{r warning=FALSE, message=FALSE}
set.seed(1)
lo$sisalCorners(tau=2^-6, nonNeg=T, iters=300)
lo$selectGenes(100)
lo$projectionPlot()
```

## Deconvolution

Once we selected signature genes, we can use them as an input for DSA algorithm to perform deconvolution.

```{r warning=FALSE, message=FALSE}
set.seed(1)
lo$deconvolve(dataset="raw")

data("proportionsLiverBrainLung")
actualProportions <- proportionsLiverBrainLung[, 10:42]

plotProportions(lo$proportions, actualProportions[c(2, 1, 3), ],
                pnames=c("Linseed", "Actual proportions"))

```
