---
title: "Linseed Tutorial"
author: "Konstantin Zaitsev"
date: "`r Sys.Date()`"
output: html_document
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---
# Linseed tutorial

Linseed (LINear Subspace identification for gene Expresion Deconvolution) is a package that provides tools and interface to explore gene expression datasets in linear space.

## Installing the package

You can install the package using `devtools::install_github`:

```{r eval=FALSE}
devtools::install_github("ctlab/linseed")
```

Current build was tested using rhub: 

## Getting started with linseed

To start working with gene expression data, we need to create a new LinseedObject, in this tutorial we will use GSE19830 (mixture of Liver, Brain and Lung), we will take only mixed samples (10-42) and will take only 10000 most expressed genes.

```{r warning=FALSE, message=FALSE}
library(linseed)
lo <- LinseedObject$new("GSE19830", samples=10:42, topGenes=10000)
```

## Coolinearity networks

To build a coolinearity network we first have to evaluate all pairwise collinearity coefficients, all pairwise spearman correlation and then run significance test which will calculate p value for each each by shuffling network weights randomly.

```{r networks, warning=FALSE}
lo$calculatePairwiseLinearity()
lo$calculateSpearmanCorrelation()
lo$calculateSignificanceLevel(100)
lo$significancePlot(0.01)

lo$filterDatasetByPval(0.01)
lo$svdPlot()

```

To visualiaze what left after filtering we can call projection plot from our object. But we have to project the data to the simplex first.

```{r visi, warning=FALSE, message=FALSE}
lo$setCellTypeNumber(3)
lo$project("full") # projecting full dataset
lo$projectionPlot(color="filtered")
```


## Deconvolution

To deconvolve the dataset, you first have to project (full or filtered dataset) to the simplex, and then find corners of it.

```{r deconvolution}
lo$project("filtered")
lo$smartSearchCorners(dataset="filtered", error="norm")
lo$deconvolveByEndpoints()
plotProportions(lo$proportions)
```

We can also use tSNE to haave an idea of how data looks like when dimensionally reduced.

```{r rtsne}
# lets select 100 genes closest to 
lo$selectGenes(100)
lo$tsnePlot()

```

To compare with actual proportions you can use `dotPlotProportions` function

```{r proportions}
data("proportionsLiverBrainLung")
dotPlotPropotions(lo$proportions, proportionsLiverBrainLung[, 10:42], guess=TRUE)

```
